---
title: "FinalProject"
output: html_document
date: "2022-11-26"
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
cb_palette = c("#999999", "#E69F00", "#56B4E9", "#009E73",
    "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
#load("lattice.RData")
```


Load some packages:
```{r}
library(dplyr)
library(tidyverse)
library(vcd)
library(ggmosaic)
library(leaps)
library(broom)
library(MASS)
library(ggplot2)
library(corrplot)
library("ggplot2")                     
library("GGally")
```
Dataset: 
We have a dataset of 10000 rows with 14 attributes to perform our analysis and give recommendations to stem attrition. 


```{r}
data <- read.csv("ChurnModeling.csv")

```

```{r}
data$Gender <- as.factor(data$Gender)
data$Geography <- as.factor(data$Geography)
data$HasChckng <- as.factor(data$HasChckng)
data$IsActiveMember <- as.factor(data$IsActiveMember)
summary(data)
```
Research Questions:
We try to address this issue by performing exploratory analysis and answer the following questions:

1)	Are all features important for determining client will churn or not?
2)	How do various features impact the likelihood of churn?
3)	Could we well describe the notions of churn using a simple logistic regression or do we need more complex models?

Correlation between Numerical Values:

```{r}
cor_matrix <- subset(data, select = -c(RowNumber,CustomerId))
cor_matrix<- cor_matrix %>% select_if(is.numeric) 

ggpairs(cor_matrix)+theme_bw()

# , upper = list(continuous = "density", combo = "box_no_facet"),lower = list(continuous = "points", combo = "dot_no_facet")
```

Credit Score
```{r}
ggplot(data, aes(x= CreditScore, fill=factor(Exited), color=factor(Exited))) +
geom_histogram(position="identity", bins = 10)+
  scale_fill_manual(values = cb_palette)
```


```{r}
ggplot( aes(x=Exited, y= CreditScore, fill= factor(Exited) ), data = data) +
    geom_boxplot() +
    ggtitle("A boxplot with Credit Score") +
    xlab("Exited")+
  scale_fill_manual(values = cb_palette)
```
Geography 
```{r}
data %>% group_by(Geography, Exited) %>% summarise(Percent = round(n()/nrow(.),2)) %>% ggplot(aes(x = Geography, y = Percent, fill = factor(Exited))) + geom_bar(stat = "identity") + geom_text(aes(label = Percent * 100), vjust = 1.5, hjust = 0.5,color = "white", size = 5.0) + theme_minimal() +
  scale_fill_manual(values = cb_palette) + scale_y_continuous(labels = scales::percent_format(accuracy = 1))
```


- We see that churn is higher in the east region.
- Central and West are doing okay

Gender

    
```{r}
data %>% filter(Gender == "Neutral")
```

```{r}

data$Gender[data$RowNumber == 17] <- "Female"
data$Gender[data$RowNumber == 34] <- "Female"
data$Gender[data$RowNumber == 226] <- "Female"
data$Gender[data$RowNumber == 380] <- "Female"
data$Gender[data$Gender == "Neutral"] <- "Male"
summary(data)
```

```{r}
data %>% group_by(Gender, Exited) %>% summarise(Percent = round(n()/nrow(.),2)) %>% ggplot(aes(x = Gender, y = Percent, fill = factor(Exited))) + geom_bar(stat = "identity") + geom_text(aes(label = Percent * 100), vjust = 1.5, hjust = 0.5,color = "white", size = 5.0) + theme_minimal() +
  scale_fill_manual(values = cb_palette) + scale_y_continuous(labels = scales::percent_format(accuracy = 1))
```

- The churn rate is higher in female clients than in male clients.
- The proportion of male clients retained is higher than the female clients.
- The proportion of Neutral gender is equal

Age
Looks like we have outliers in age 

```{r}
ggplot(data, aes(x = factor(Exited), y = Age, fill = factor(Exited))) + 
  geom_boxplot() +
  stat_summary(fun = "mean", geom = "point", shape = 8,
               size = 2, color = "black")+
  scale_fill_manual(values = cb_palette)
  
  
  
```
1. We can see some noise in the Age column for the clients who are retained. That is age cannot be 190, there is one other client with age 101.
2. There is literally one client with 30y tenure and we see that as an outlier in our plot.

Let's deepdive in to the outliers 

```{r}
ceiling(mean(data$Age[data$Gender == "Male" & data$Geography == "West"]))
data$Age[data$Age > 100] <- ceiling(mean(data$Age[data$Gender == "Male" & data$Geography == "West"]))
summary(data)
```

Tenure

```{r}
data %>% group_by(Tenure, Exited) %>% summarise(Percent = round(n()/nrow(.),2)) %>% ggplot(aes(x = Tenure, y = Percent, fill = factor(Exited))) + geom_bar(stat = "identity") + geom_text(aes(label = Percent * 100), vjust = 1.5, hjust = 0.5,color = "white", size = 5.0) + theme_minimal() +
  scale_fill_manual(values = cb_palette) + scale_y_continuous(labels = scales::percent_format(accuracy = 1))
```

```{r}
ggplot(data, aes(x = factor(Exited), y = Tenure, fill = factor(Exited))) + 
  geom_boxplot() +
  stat_summary(fun = "mean", geom = "point", shape = 8,
               size = 2, color = "black")+
  scale_fill_manual(values = cb_palette)
  
  
  
```
Balance
```{r}
ggplot(data, aes(x = factor(Exited), y = Balance, fill = factor(Exited))) + 
  geom_boxplot() +
  stat_summary(fun = "mean", geom = "point", shape = 8,
               size = 2, color = "black")+
  scale_fill_manual(values = cb_palette)
  
  
  
```
NumOfProducts

```{r}
data %>% group_by(NumOfProducts, Exited) %>% summarise(Percent = round(n()/nrow(.),2)) %>% ggplot(aes(x = NumOfProducts, y = Percent, fill = factor(Exited))) + geom_bar(stat = "identity") + geom_text(aes(label = Percent * 100), vjust = 1.5, hjust = 0.5,color = "white", size = 5.0) + theme_minimal() +
  scale_fill_manual(values = cb_palette) + scale_y_continuous(labels = scales::percent_format(accuracy = 1))
```
EstimatedSalary

```{r}
ggplot(data, aes(x = factor(Exited), y = EstimatedSalary, fill = factor(Exited))) + 
  geom_boxplot() +
  stat_summary(fun = "mean", geom = "point", shape = 8, size = 2, color = "black")+
  scale_fill_manual(values = cb_palette)

ggplot(data, aes(x=EstimatedSalary)) + 
  geom_histogram()
```

```{r}

fit <- glm(formula = Exited ~ CreditScore + Geography + Gender + Age + Tenure+ Balance + NumOfProducts + HasChckng + IsActiveMember + EstimatedSalary, family = binomial, data = data)
summary(fit)
```
Significant variables, for this specific model, are those with a p-value less than 0.05.

```{r}
coefs <- tidy(fit, conf.int = TRUE)
ggplot(coefs[-1, ], aes(x = estimate, y = term, xmin = conf.low, xmax = conf.high)) +
  geom_point() + geom_errorbarh() + geom_vline(xintercept = 0) + labs(y="Features") + ggtitle("Confidence interval of estimates for each term")

```




```{r}
ggplot(data = data, aes(x = NumOfProducts, y = Age, color= factor(Exited), shape= factor(Exited))) + 
  geom_point() +
  facet_wrap(~Geography) 
  
```
```{r}
ggplot(data = data, aes(x = NumOfProducts, y = Age, color= factor(Exited), shape= factor(Exited))) + 
  geom_point() +
  facet_wrap(~IsActiveMember) 
  
```
```{r}
ggplot(data = data, aes(x = NumOfProducts, y = Age, color= factor(Exited), shape= factor(Exited))) + 
  geom_point() +
  facet_wrap(~Gender) 
  
```
```{r}
ggplot(data = data, aes(x = Balance, y = Age, color=factor(Exited), shape= factor(Exited))) + 
  geom_point() +
  facet_wrap(~Gender)
```







```{r}
# Age
ggplot(data, aes(x = Age, y = Exited, color = factor(Exited))) +
  geom_jitter(width = 0, height = 0.1) + ggtitle("Likelihood of Churn")+ 
  scale_color_manual(name="Churn",values = c("0"="darkorange","1"="deepskyblue1")) 
#+  scale_y_continuous(breaks=c(0,1))





```

```{r}
ggplot(data, aes(x = Age, fill = factor(Exited))) +
  geom_histogram(breaks = seq(25, 75, 5), position = 'fill') +
  ggtitle("Likelihood of churn")+ 
  guides(fill=guide_legend(title="Churn")) + ylab('Probability')
```

```{r}


gg = ggplot(data, aes(x = Age, y = Exited)) +
  geom_jitter(width = 0, height = 0.1) +
  geom_smooth(method = "loess",aes(color='loess'))

gg + geom_smooth(method = "glm", method.args = list(family = "binomial"), aes(color = "logistic")) + 
  scale_color_manual(name="Fit",values = c("loess"="darkorange","logistic"="deepskyblue1"))+   ggtitle("Loess and logistic Fit")

```
```{r}


gg = ggplot(data, aes(x = Balance, y = Exited)) +
  geom_jitter(width = 0, height = 0.1) +
  geom_smooth(method = "loess",aes(color='loess'))

gg + geom_smooth(method = "glm", method.args = list(family = "binomial"), aes(color = "logistic")) + 
  scale_color_manual(name="Fit",values = c("loess"="darkorange","logistic"="deepskyblue1"))+   ggtitle("Loess and logistic Fit")

```

```{r}
data.logit = glm(Exited ~ Age + NumOfProducts, family = "binomial", data = data)

summary(data.logit)

```
```{r}

grid_df = expand.grid(Age = 5:95, NumOfProducts = seq(1, 4, 1))
grid_preds = augment(data.logit, type.predict = "response", newdata = grid_df)
ggplot(grid_preds, aes(x = Age, y = .fitted, group = factor(NumOfProducts), color = factor(NumOfProducts))) +
  geom_line() +
  xlab("Age") +
  ylab("Probability of churn") +
  labs(color = "NumOfProducts")+
  scale_color_manual(values = cb_palette)+
  ggtitle("Age vs Probability of churn for each values of NumOfProducts") 


```
```{r}
grid_df2 = expand.grid(Age = seq(5, 95, 5), NumOfProducts = seq(1, 4, 1))
grid_preds = augment(data.logit, type.predict = "response", newdata = grid_df2)
ggplot(grid_preds, aes(x = NumOfProducts, y = .fitted, group = Age, color = Age)) +
  geom_line() +
  xlab("NumberOfProducts") +
  ylab("Probability of churn") +
  labs(color = "Age")+  
  
  #scale_x_continuous(breaks=c(-3,0,3,6,9)) +
  scale_color_viridis_c()+
  ggtitle("NumberOfProducts vs Probability of churn for each groups of Age")
```


```{r}
data.logit = glm(Exited ~ Age + Balance, family = "binomial", data = data)

summary(data.logit)

```

```{r}

grid_df = expand.grid(Age = 5:95, Balance = seq(0,250898, 1000))
grid_preds = augment(data.logit, type.predict = "response", newdata = grid_df)
ggplot(grid_preds, aes(x = Age, y = .fitted, group = Balance, color = Balance)) +
  geom_line() +
  xlab("Age") +
  ylab("Probability of churn") +
  labs(color = "Balance")+
  scale_color_viridis_c()+
  ggtitle("Age vs Probability of churn for each values of Balance") 


```

```{r}
grid_df2 = expand.grid(Age = seq(5, 95, 5), Balance = seq(0, 250898, 1000))
grid_preds = augment(data.logit, type.predict = "response", newdata = grid_df2)
ggplot(grid_preds, aes(x = Balance, y = .fitted, group = Age, color = Age)) +
  geom_line() +
  xlab("Balance") +
  ylab("Probability of churn") +
  labs(color = "Age")+  
  scale_color_viridis_c()+
  ggtitle("Balance vs Probability of churn for each groups of Age")
```

```{r}
data
```


```{r}
data.logit = glm(Exited ~ NumOfProducts + Balance, family = "binomial", data = data)

summary(data.logit)

```


```{r}

grid_df = expand.grid(NumOfProducts = c(1,2,3,4), Balance = seq(0,250898, 1000))
grid_preds = augment(data.logit, type.predict = "response", newdata = grid_df)
ggplot(grid_preds, aes(x = NumOfProducts, y = .fitted, group = Balance, color = Balance)) +
  geom_line() +
  xlab("NumOfProducts") +
  ylab("Probability of churn") +
  labs(color = "Balance")+
  scale_color_viridis_c()+
  ggtitle("NumOfProducts vs Probability of churn for each values of Balance") 


```



```{r}
grid_df2 = expand.grid(NumOfProducts = c(1,2,3,4), Balance = seq(0, 250898, 1000))
grid_preds = augment(data.logit, type.predict = "response", newdata = grid_df2)
ggplot(grid_preds, aes(x = Balance, y = .fitted, group = NumOfProducts, color = NumOfProducts)) +
  geom_line() +
  xlab("Balance") +
  ylab("Probability of churn") +
  labs(color = "Age")+  
  scale_color_viridis_c()+
  ggtitle("Balance vs Probability of churn for each groups of NumOfProducts")
```




```{r}
data.logit.lot <- glm(Exited ~ Age + Balance + NumOfProducts + IsActiveMember + Gender+ Geography, family = "binomial", data =data)
summary(data.logit.lot)

data_model_df = augment(data.logit.lot, type.residuals = "pearson")
ggplot(data_model_df, aes(x = .fitted, y = .resid)) +
  geom_point() +
  geom_smooth(method = "loess", method.args = list(degree = 1)) +
  xlab("Fitted values") +
  ylab("Residuals") + ggtitle("Residual plot of FittedValues vs Residuals")

ggplot(data_model_df, aes(x = Age, y = .resid)) +
  geom_point() +
  geom_smooth(method = "loess", method.args = list(degree = 1)) +
  xlab("Age") +
  ylab("Residuals")

ggplot(data_model_df, aes(x = NumOfProducts, y = .resid)) +
  geom_point() +
  geom_smooth(method = "loess", method.args = list(degree = 1)) +
  xlab("NumOfProducts") +
  ylab("Residuals")
```

